\section{Illustrative Examples}

We present examples in two groups. The first group consists of two analytically tractable cases where the inference problem is provably non-identifiable; these require no numerical computation and serve to illustrate the theoretical conditions of \Cref{section identifiability analysis}. The second group consists of two computational examples where all identifiability conditions are satisfied. For the numerical examples, we compare our Bayesian MCMC approach against the sparse-learning-CRN method of \cite{zhang2019learning}, which recovers propensity functions by solving an $\ell_1$-regularized least-squares problem (LASSO) over the polynomial basis $\{1,A,B,A(A{-}1)/2,AB,B(B{-}1)/2\}$ using the FISTA algorithm. (All code and data are available at \href{https://github.com/ssindiUCM/BayesCRNInference}{BayesCRNInference}.)

%%--------------------------------------------------------------------
\subsection*{Structural Non-Identifiability: Two Illustrative Cases}
%%--------------------------------------------------------------------

The following two cases both satisfy \Cref{corollary structural non-identifiability by topological conditions} via a non-negative conservation law. In each case the violation of condition~\eqref{eq. sufficient and necessary conditions for structural identifiability} persists for all trajectory lengths: no amount of data can resolve the ambiguity. We do not report MCMC results for these cases; the non-identifiability is established analytically, and running the sampler would merely confirm that the posterior is diffuse over an equivalence class of parameter vectors.

\paragraph{First case: stoichiometric conservation law.} Following \cite{enciso2021identifiability}, consider two systems on species $(X_1,X_2)$ sharing the conservation law $X_1+X_2=2$:
\begin{equation}
X_1\xrightarrow{k_1=1}X_2,\quad X_2\xrightarrow{k_2=1}X_1
\label{eq:deg1}
\end{equation}
\begin{equation}
2X_1\xrightarrow{\lambda_1=2}X_1+X_2,\quad X_1+X_2\xrightarrow{\lambda_2=1}2X_1,\quad
X_1+X_2\xrightarrow{\lambda_3=1}2X_2,\quad 2X_2\xrightarrow{\lambda_4=2}X_1+X_2
\label{eq:deg2}
\end{equation}
The conservation law confines the state space to $\{(2,0),(1,1),(0,2)\}$. Table~\ref{tab:degenerate} shows that both systems produce identical propensities at every accessible state, so condition~\eqref{eq. sufficient and necessary conditions for structural identifiability} is violated for every member of the interpolating family $\theta(\alpha)=\alpha\,\theta_{\text{Sys.1}}+(1-\alpha)\,\theta_{\text{Sys.2}}$. The conservation vector $c=(1,1)^\top$ exists and the channel $\xi_\ell=(-1,+1)$ contains the unimolecular reaction $X_1\to X_2$ ($\|\nu\|_1=1$), satisfying condition~(2) of \Cref{corollary structural non-identifiability by topological conditions}. Structural non-identifiability follows regardless of trajectory length.

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
State $(X_1,X_2)$ & $(\Delta X_1,\Delta X_2)$ & Sys.~\eqref{eq:deg1} propensity & Sys.~\eqref{eq:deg2} propensity \\
\hline
$(2,0)$ & $(-1,+1)$ & $k_1\cdot 2=2$ & $\lambda_1\cdot\tfrac{2(2-1)}{2}=2$ \\
$(1,1)$ & $(-1,+1)$ & $k_1\cdot 1=1$ & $\lambda_3\cdot 1\cdot 1=1$ \\
$(1,1)$ & $(+1,-1)$ & $k_2\cdot 1=1$ & $\lambda_2\cdot 1\cdot 1=1$ \\
$(0,2)$ & $(+1,-1)$ & $k_2\cdot 2=2$ & $\lambda_4\cdot\tfrac{2(2-1)}{2}=2$ \\
\hline
\end{tabular}
\caption{Propensities for Systems~\eqref{eq:deg1} and~\eqref{eq:deg2} with $X_1+X_2=2$. The two systems are indistinguishable at every accessible state.}
\label{tab:degenerate}
\end{table}

\paragraph{Second case: catalyst species.} Consider the two-species system
\begin{equation}
\emptyset\xrightarrow{k_1=1}X_1,\quad X_1\xrightarrow{k_2=1}2X_1,\quad X_1+X_2\xrightarrow{k_3=0.1}X_2,
\label{eq:example2}
\end{equation}
with propensities $a_1=k_1$, $a_2=k_2 X_1(X_1{-}1)/2$, $a_3=k_3 X_1 X_2$. Since no reaction in~\eqref{eq:example2} produces or degrades $X_2$, the species acts as a permanent catalyst: $X_2(t)=c_2$ for all $t$. This constitutes a conservation law with $c=(0,1)^\top$, and \Cref{corollary structural non-identifiability by topological conditions} applies again. In the channel $\xi_\ell=(-1,0)$, the candidate reaction $X_1\to\emptyset$ is unimolecular ($\|\nu\|_1=1$), satisfying condition~(2). The propensities $k\,X_1$ (for $X_1\to\emptyset$) and $k'\,c_2 X_1$ (for $X_1+X_2\to X_2$) are proportional for all visited states, so the full-rank condition~\eqref{eq. local cancavity check} fails and any pair $(\tilde k_1,\tilde k_3)$ with $\tilde k_1+\tilde k_3 c_2=k_3 c_2$ is equally consistent with the data. In contrast to the first case, where non-identifiability arises from the topology of the state space, here it arises from collinearity of the propensity functions - but both are structural, persisting in every trajectory generated by these networks.

\subsubsection*{Numerical Example 1: Predator-Prey System}

Both numerical examples satisfy all identifiability conditions of \Cref{section identifiability analysis}. Throughout, Bayesian MCMC is run with a spike-and-slab prior ($\pi=0.75$, $\alpha=2$, $\beta=0.25$), 500{,}000 iterations, 50{,}000 burn-in, thinning by 100, and rates initialized from $U[0.001,2.0]$, with covariance updated adaptively every 10 iterates. For comparison, the LASSO method of \cite{zhang2019learning} is applied at regularization strength $\lambda=0.01$ unless stated otherwise.

We consider the stochastic Lotka-Volterra system with prey $A$ and predator $B$:
\begin{equation}
A\xrightarrow{k_1=1.0}2A,\quad
A\xrightarrow{k_2=0.1}\emptyset,\quad
A+B\xrightarrow{k_3=0.01}2B,\quad
B\xrightarrow{k_4=0.5}\emptyset,
\label{eq:predprey}
\end{equation}
with mass-action propensities $\lambda_1=k_1 A$, $\lambda_2=k_2 A$, $\lambda_3=k_3 AB$, $\lambda_4=k_4 B$. The deterministic equilibrium is $(A^*,B^*)=(50,90)$. Starting from $(A,B)=(100,50)$, the system is simulated over $T=40$, yielding 5{,}805 events across four stoichiometric channels. The large initial displacement from equilibrium drives pronounced oscillatory dynamics, with stochastic prey extinction occurring near $t=39.8$.

We search over all unimolecular and bimolecular candidate reactions for two species. The rank condition~\eqref{eq. local cancavity check} is satisfied for each active channel. Table~\ref{tab:predprey_detail} shows all candidate reactions considered for each stoichiometric channel, together with the posterior results.

\input{table_predprey_detail}

The Bayesian approach correctly identifies all four stoichiometric channels, with posterior means within 1\% of the true rates and $\Pr(\text{off})=0$ for all active reactions. All inactive candidate reactions receive $\Pr(\text{off})\approx 1$.

The LASSO performs well on channels $(+1,0)$, $(-1,+1)$, and $(0,-1)$, but fails on channel $(-1,0)$ corresponding to $A\to\emptyset$: rather than recovering a propensity proportional to $A$, it returns $0.018\,A+0.036\,B$, assigning the larger coefficient to $B$. This is mechanistically impossible under mass-action kinetics. The failure has a structural cause: near the deterministic equilibrium $B^*=90>A^*=50$, values of $B$ are systematically larger than values of $A$, so assigning weight to $B$ achieves a similar propensity contribution at lower $\ell_1$ cost. The oscillation-induced correlation between $A$ and $B$ further compounds this, and the low event count (190 jumps) provides insufficient signal to overcome it. Because the Bayesian approach works directly in the space of reaction rates under mass-action constraints — rather than fitting unconstrained polynomial coefficients — it is not susceptible to this scale imbalance.

\subsubsection*{Numerical Example 2: Randomly Generated Reaction Network}

Using the same MCMC and LASSO settings as in the previous example, we assess performance at scale with a four-species system $(A,B,X,Y)$, enumerating all unimolecular and bimolecular reactions to form a candidate space of 130 stoichiometric channels. Twenty reactions are selected at random (with 4 channels containing 2 active reactions each), and rate parameters are drawn from $\mathrm{Gamma}(2.6,\,0.4)$. The system is simulated from $(A,B,X,Y)=(5,5,5,5)$ over $T\in\{200,400\}$ on a single trajectory. MCMC is run independently on each of the 15 channels with observed jumps.

Results in Tables~\ref{tab:recovery_combined}--\ref{tab:lasso-lambda-sweep} show that the spike-and-slab prior achieves exact structural recovery at both time horizons: every inactive reaction receives $\Pr(\text{off})=1.000$ and every active reaction $\Pr(\text{off})=0.000$, with posterior widths narrowing as $T$ increases. At $T=100$, the Bayesian method achieves lower $L^2$ propensity error than LASSO on every channel (errors $0.008$--$0.168$ vs.\ $0.143$--$0.991$). A sweep over $\lambda\in\{0.001,0.005,0.01,0.05,0.1,0.5\}$ confirms that no single regularization strength minimizes error across all channels simultaneously: weak-reaction channels require small $\lambda$ to avoid over-shrinkage, while noisy channels require large $\lambda$ for suppression. The Bayesian framework resolves this automatically through the spike-and-slab prior, adapting to each channel's evidence without global tuning.

% True system description
\input{table1_true_system}

% Bayesian detail at T=100, T=200, T=400
\input{table3b_bayes_detail_T100}
\input{table3b_bayes_detail_T200}
\input{table3b_bayes_detail_T400}

% L2 comparison: Bayes vs LASSO
\input{table3_comparison_l2}

% LASSO fixed lambda across T
\input{table2a_lasso_fixed_lambda}

% LASSO lambda sweep at T=200
\input{table2b_lasso_lambda_sweep}
